---
title: "Learnr exercises in penguinsbox"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Types of exercises of learnr available in penguinsbox.
---

<!-- Add JavaScript code for making the exercise code larger -->
<script language="JavaScript" src="js/exercise-font-size.js"></script>

```{r setup, include=FALSE}
# load packages-----------------------------------------------------------------

library(learnr)
library(parsons)
library(tidyverse)
library(palmerpenguins)


# set options for exercises and checking ---------------------------------------

# hide non-exercise code chunks ------------------------------------------------
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction: learning learnr

<iframe src="https://giphy.com/embed/l0E9CrSVVI3g4" width="480" height="204" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/movie-film-cute-l0E9CrSVVI3g4">via GIPHY</a></p>

Let's take the first penguin steps into [learnr](https://github.com/rstudio/learnr) exercises.

### Learning goals

Show you the types of exercises available in `penguinsbox`.


## Filling the blanks

Along the exercises of `penguinsbox` package we will be using two main packages the [tidyverse](https://www.tidyverse.org/) metapackage for data science and [palmerpenguins](https://github.com/allisonhorst/palmerpenguins/) package which contains the penguins dataset. We can load the two packages below (first tidyverse, and then palmerpenguins) by replacing the `___` with the package names. You can type directly on to of `___` in the box below; if you do not know the answer, you can click `Solution`, click `Copy to Clipboard`, paste it in the box and then click `Submit Answer`.

```{r library-tidyverse, exercise=TRUE}
# load the tidyverse and palmerpenguins packages
library(___)
library(___)
```

```{r library-tidyverse-solution}
# load the tidyverse and palmerpenguins packages
library(tidyverse)
library(palmerpenguins)
```

```{r library-tidyverse-check}
grade_code("You completed the first filling the blanks exercise.")
```

## MCQ

In these exercises we will be presented with Multiple choice questions (MCQ) in multiple quizes. Some of the questions may require multiple answers like the one below. Tick the correct answers. 

```{r sentiment-quiz}
question('What are the first and the last letters of the word Penguins? (select BOTH first and last lettes)',
         answer("P", correct = TRUE),
         answer("s", correct = TRUE),
         answer("e"),
         answer("n"),
         allow_retry = TRUE,
         random_answer_order = TRUE,
         incorrect = "Are you sure these are the first and last letters of the word Penguins?",
         correct = "Correct! P and s first and last letters of the word Penguins respectively"
)
```

## Parsons problems

Some of the exercises will be Parsons problems. In a Parsons problem you will be presented with the solution, but the lines of code will be jumbled. The task is to arrange these lines in the correct order. 
For example, to write `Hi penguins!` we need to drag and drop the words from the panel on the left to the panel on the right in the correct order. 

```{r explain-1, echo =FALSE}
question_parsons(
  initial = c(
 "penguins!",
    "Hi"
  ),
  answer(
    c(
    "Hi",
    "penguins!"
    ),
    correct = TRUE
  ),
  problem_type = "base"
)
```


## Wrap up

In `penguinsbox` package we will encounter three types of exercises:

- Filling in the blanks or fading exercises. 
- Multiple choice questions.
- Parsons problems.