---
title: "Factors"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Use forcats functions to manipulate factors in R.
---

<!-- Add JavaScript code for making the exercise code larger -->
<script language="JavaScript" src="js/exercise-font-size.js"></script>

```{r setup, include=FALSE}
# load packages-----------------------------------------------------------------

library(learnr)
library(tidyverse)
library(palmerpenguins)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = gradethis::grade_learnr,warnings= FALSE)
# set options for exercises and checking ---------------------------------------


# hide non-exercise code chunks ------------------------------------------------
tutorial_options(exercise.timelimit = 60)
```


## Introduction

<iframe src="https://giphy.com/embed/cGzafVX6A8ewSRruJA" width="480" height="270" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/natgeochannel-nat-geo-national-geographic-hostile-planet-cGzafVX6A8ewSRruJA">via GIPHY</a></p>


This chapter will focus on the **forcats** package for categorical variables manipulation, which is part of the core tidyverse.


### Learning objectives

In this tutorial you are going to practice some forcats functions that let you:

* Inspect the levels of a factor.
* Display the levels of a factor as they are. 
* Reorder the levels.
* Collapse the levels.

###


## Find levels
```{r test-factor, echo = FALSE}
question("How could you determine the levels of species factor? Select all options that apply",
         answer("factors(penguins$species)", message="Close! But there is a typo in this answer"),
         answer("levels(penguins$species)", correct = TRUE),
         answer("levels(factor(penguins$species))", correct = TRUE),
         allow_retry = TRUE )
```

###
As we see the order of the levels does not need to match the order in which the data is passed in a dataframe. 

```{r, echo=TRUE, warning = FALSE}
ggplot(penguins, aes(species,body_mass_g)) +
    geom_boxplot()
```

## Reorder factors

Now that we know which are the levels of `species`, we can start reordering them. For example, lets make a geom_boxplot to plot `body_mass_g` for each species

```{r, echo=TRUE}
ggplot(penguins, aes(species, body_mass_g)) +
    geom_boxplot()
```

###

As we see the factors are plotted in the order of the levels. What if we want to plot in the order of appeared?

```{r str-fct-inorder, exercise=TRUE}
ggplot(penguins, aes(___(species), body_mass_g)) +
    geom_boxplot()
```

```{r str-fct-inorder-hint-1}
Which function of forcats could reorder factor levels by first appearance?
```


```{r str-fct-inorder-solution}
ggplot(penguins, aes(fct_inorder(species), body_mass_g)) +
    geom_boxplot()
```

```{r str-fct-inorder-check}
grade_code("fct_inorder reorders factor levels by order of appearance.")
```

###

We can also reorder the species based on `body_mass_g` values:

```{r str-fct-reorder, exercise=TRUE}
ggplot(penguins, aes(___(___, ___),___)) +
    geom_boxplot()
```

```{r str-fct-reorder-hint-1}
ggplot(penguins, aes(fct_reorder(___, ___),___)) +
    geom_boxplot()
```

```{r str-fct-reorder-hint-2}
Check `??forcats::fct_reorder`
```

```{r str-fct-reorder-solution}
ggplot(penguins, aes(fct_reorder(species, body_mass_g),body_mass_g)) +
    geom_boxplot()
```


## Collapse factors

Next thing we are going to do is lumping together factors. Let's see how many penguins of each specie there are in the palmerpenguins dataset:

```{r penguins-species, exercise = TRUE}
penguins %>%
  count(species)
```

###

```{r penguins-species-solution}
penguins %>%
  count(species)
```

###

We could retain the specie with more number of penguins (i.e. Adelie) and lump together the other species into a level called "other".  

```{r penguins-species-fctlump, exercise = TRUE}
penguins %>%
  mutate(species = fct_lump(species,1)) %>% 
  count(species) 
```

```{r penguins-species-fctlump-solution}
penguins %>%
  mutate(species = fct_lump(species,1)) %>% 
  count(species) 
```

###

We can also lump into level other levels that levels that appear fewer than min times like this:

```{r penguins-species-fctlump-min, exercise = TRUE}
penguins %>%
  mutate(species = fct_lump_min(species,min=70, other_level = "less_70")) %>% 
  count(species) 
```

```{r penguins-species-fctlump-min-solution}
penguins %>%
  mutate(species = fct_lump_min(species,min=70, other_level = "less_70")) %>% 
  count(species) 
```

```{r penguins-species-fctlump-min-check}
grade_code("Species with less than 70 penguins are grouped to less_70 level")
```


## Relevel factors

We can manually reorder factor levels using `fct_relevel`. For example, we can move Gentoo to be the first level position in species.  

```{r, echo =TRUE}
penguins %>%
    mutate(species= fct_relevel(species, "Gentoo", after=0)) %>% 
    ggplot(aes(species, body_mass_g)) +
    geom_boxplot()
```

## Wrap up 

You have practice some of the tools for working with categorical variables (Factors).  



