---
title: "Factors"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Use forcats functions to manipulate factors in R.
---

<!-- Add JavaScript code for making the exercise code larger -->
<script language="JavaScript" src="js/exercise-font-size.js"></script>

```{r setup, include=FALSE}
# load packages-----------------------------------------------------------------

library(learnr)
library(tidyverse)
library(palmerpenguins)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = gradethis::grade_learnr)

knitr::opts_chunk$set(echo = FALSE, exercise.checker = gradethis::grade_learnr)
# set options for exercises and checking ---------------------------------------


# hide non-exercise code chunks ------------------------------------------------
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

<iframe src="https://giphy.com/embed/cGzafVX6A8ewSRruJA" width="480" height="270" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/natgeochannel-nat-geo-national-geographic-hostile-planet-cGzafVX6A8ewSRruJA">via GIPHY</a></p>


This chapter will focus on the **forcats** package for categorical variables manipulation, which is part of the core tidyverse.


### Learning objectives

In this tutorial you are going to practice some forcats functions that let you:


* Inspect the levels of a factor.
* Display the levels of a factor as they are. 
* Reorder the levels.
* Collapse the levels.

###


## Find levels
```{r test-factor, echo = FALSE}
question("How could you determine the levels of species factor? Select all options that apply",
         answer("factors(penguins$species)", message="Close! But there is a typo in this answer"),
         answer("levels(penguins$species)", correct = TRUE),
         answer("levels(factor(penguins$species))", correct = TRUE),
         allow_retry = TRUE )
```

###
As we see the order of the levels does not need to match the order in which the data is passed in a dataframe. 

```{r}
ggplot(penguins, aes(species,body_mass_g)) +
    geom_boxplot()
```


## Reorder factors

Now that we know which are the levels, we can start reordering them. For example, lets make a geom_boxplot to plot `body_mass_g` for each species

```{r}
ggplot(penguins, aes(species, body_mass_g)) +
    geom_boxplot()
```

As we see the factors are plotted in the order of the levels. What if we want to plot in the order of appeared?

```{r str-fct-inorder, exercise=TRUE}
ggplot(penguins, aes(___(species), body_mass_g)) +
    geom_boxplot()
```


```{r str-fct-inorder-solution}
ggplot(penguins, aes(fct_inorder(species), body_mass_g)) +
    geom_boxplot()
```

###

Fantastic! We can also reorder the species based on body_mass_g values

```{r str-fct-reorder, exercise=TRUE}
ggplot(penguins, aes(___(species, body_mass_g),body_mass_g)) +
    geom_boxplot()
```


```{r str-fct-reorder-solution}
ggplot(penguins, aes(fct_reorder(species, body_mass_g),body_mass_g)) +
    geom_boxplot()
```

Awesome! As we see the mean body_mass_g for Adelie species is slighly smaller than for Chinstrap, and therefore this plot sorts Adelie specie gets sorted.


## Collapse factors

```{r}
penguins %>%
  count(species)
```



```{r}
penguins %>%
  mutate(species = fct_lump(species,1)) %>% 
  count(species) 
```
Other will be a factor...


```{r}
penguins %>%
  mutate(species = fct_lump(species,1)) %>% 
ggplot(aes(species, body_mass_g)) +
    geom_boxplot()
  
```

We can also lump to a minimum number of lumps like this 
  
```{r}
penguins %>%
  mutate(species = fct_lump_min(species,min=70)) %>% 
  ggplot(aes(species, body_mass_g)) +
    geom_boxplot()
```



## Relevel factors

We can reorder factor levels by hand using `fct_relevel`. For example, we can move Gentoo to be the first level position in species.  

```{r}
penguins %>%
    mutate(species= fct_relevel(species, "Gentoo", after=0)) %>% 
    ggplot(aes(species, body_mass_g)) +
    geom_boxplot()
```
