---
title: "Strings"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Use string functions to manipulate string character vectors in R .
---

<!-- Add JavaScript code for making the exercise code larger -->
<script language="JavaScript" src="js/exercise-font-size.js"></script>

```{r setup, include=FALSE}
# load packages-----------------------------------------------------------------

library(learnr)
library(tidyverse)
library(palmerpenguins)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = gradethis::grade_learnr)

# set options for exercises and checking ---------------------------------------


# hide non-exercise code chunks ------------------------------------------------
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

<iframe src="https://giphy.com/embed/C9AuY1VOvurxm" width="480" height="270" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/C9AuY1VOvurxm">via GIPHY</a></p>

This chapter will focus on the **stringr** package for string manipulation, which is part of the core tidyverse.
We recomend you to practice with tutorial 04-dplyr before starting with this tutorial.

### Learning objectives

This tutorial has few fading exercises for you to practice some stringr functions that let you:

* Find the length of matches.
* Extract the content of matches.
* Determine which strings match a pattern.
* Replace matches with new values


## Exercises 

### 1st exercise

Calculate the number of letters of the `species` variable of the `penguins` dataset:


```{r str-length, exercise=TRUE}
penguins %>% 
   mutate(length_species=___(species)) %>%
    group_by(species) %>% 
    summarise(length_species=unique(length_species))

```

```{r str-length-solution}
penguins %>% 
   mutate(length_species=str_length(species)) %>%
    group_by(species) %>% 
    summarise(length_species=unique(length_species))  
```

```{r str-length-check}
grade_code("You've successfully count the number of characters in `species` variable!")
```

###

### 2nd exercise

Extract the last two letters of `species` variable and add result to variable called `species_last_twoletter`, which is placed after species variable:

```{r str-sub, exercise=TRUE}
penguins %>% 
    mutate(species_last_twoletter= ___(species,start=___,end=-1), .after=species)
```

```{r str-sub-hint-1}
penguins %>% 
    mutate(species_last_twoletter= ___(species,start=-2,end=-1), .after=species)
```

```{r str-sub-solution}
penguins %>% 
    mutate(species_last_twoletter= str_sub(species,start=-2,end=-1), .after=species) 
```

```{r str-sub-check}
grade_code("Using str_sub you can extract the values of  character vector from the start position to the end position both inclusive")
```

###

### 3rd exercise

Add a column called `oo_species` that contains `oo` if the pattern `oo` is found in `species`. Then use a filter function to keep the species that contain  `oo`.

```{r str-extract, exercise=TRUE}
mutate(penguins,oo_species= ___(species,"___"), .after=species) %>% 
    filter(___(oo_especies))
```

```{r str-extract-hint-1}
mutate(penguins,oo_species= ___(species,"oo"), .after=species) %>% 
    filter(___(oo_especies))
```

```{r str-extract-hint-2}
mutate(penguins,oo_species= ___(species,"oo"), .after=species) %>% 
    filter(!is.na(oo_especies))
```

```{r str-extract-hint-3}
mutate(penguins,oo_species= str_extract(species,"oo"), .after=species) %>% 
    filter(!is.na(oo_especies))
```


```{r str-extract-solution}
mutate(penguins, oo_species= str_extract(species,"oo"), .after=species) %>% 
    filter(!is.na(oo_species))
```

```{r str-extract-check}
grade_code("As you've seen str_extract will add na if it doesn't find a match to extract")
```

###

### 4rd exercise

Add a column called `oo_species` that contains the character string `oo` if the pattern `oo` is found in species. Then use a filter function to keep the species that contain  `oo`.

```{r str-detect, exercise=TRUE}
mutate(penguins, oo_species= ___(species,"oo$"), .after=species) %>% 
filter(oo_species)
```


```{r str-detect-solution}
mutate(penguins, oo_species= str_detect(species,"oo$"), .after=species) %>% 
filter(oo_species)
```

```{r str-detect-check}
grade_code("str_detect adds FALSE where it doesn't find a match to extract.")
```





